---
title: "Reproducible Research - Peer-graded Assignment: Course Project 1"
output: html_document
keep_md: true 
---
## Read and check the data

1.Reading in the dataset 
```{r, echo = TRUE, cache = TRUE}
mydata<-read.csv("activity.csv")
```
2.Checking the data
```{r, echo = TRUE}
head(mydata, n=10)
summary(mydata)
```

## Plot the total number of steps taken each day

1.Aggregte steps by date
```{r, echo = TRUE}
TotalsPerDay <- aggregate(mydata$steps, list(mydata$date), FUN=sum)
colnames(TotalsPerDay) <- c("Date", "Steps")
head(TotalsPerDay, n=10)
```

2.Plot histogram of the total number of steps taken each day
```{r, echo = TRUE, warning = FALSE, message = FALSE}
install.packages(
   "ggplot2",
   repos = c("http://rstudio.org/_packages","http://cran.rstudio.com"),
   dependencies=TRUE
)
library(ggplot2)

g <- ggplot(TotalsPerDay, aes(Steps))
g <- g + geom_histogram(boundary=0, binwidth=1000,col="black", fill="lightblue") + 
         ggtitle("Total steps per day") + xlab("Steps") + ylab("Frequency")
print(g)
```

## Mean and median number of steps taken each day
1.Calculate and report the mean
```{r, echo = TRUE}
as.integer(mean(TotalsPerDay$Steps, na.rm=TRUE))
```

2.Calculate and report the median
```{r, echo = TRUE}
as.integer(median(TotalsPerDay$Steps, na.rm=TRUE))
```

## Time series plot of the average number of steps taken
1.Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE}
StepsByInterval <- aggregate(steps~interval,data=mydata,FUN=mean,na.action=na.omit)
head(StepsByInterval)

g <- ggplot(StepsByInterval, aes(interval, steps))
g <- g + geom_line(col="blue") + ggtitle("Average number of steps per interval")+
         xlab("5-minute interval") + ylab("Average number of steps")
print(g)
```

2.The 5-minute interval, on average across all the days,that contains the maximum number of steps

```{r, echo = TRUE}
StepsByInterval <- aggregate(steps~interval,data=mydata,FUN=mean,na.action=na.omit)
head(StepsByInterval)

StepsByInterval[which.max(StepsByInterval$steps), ]$interval
```

## Imputing missing data

1.Calculate and report the number of missing values
```{r, echo = TRUE}
sum(is.na(mydata))
```

2.Filling the missing (NA) values with mean values for intervals
```{r, echo = TRUE}
ImputtedData <- mydata
na <- is.na(ImputtedData$steps)
mydata_nona <- subset(mydata, !is.na(mydata$steps))
meanval <- tapply(mydata_nona$steps, mydata_nona$interval, mean, na.rm=TRUE, simplify=T)
ImputtedData$steps[na] <- meanval[as.character(ImputtedData$interval[na])]
```
3.Check the new dataset
```{r, echo = TRUE}
head(ImputtedData)
sum(is.na(ImputtedData))
```

4.Plot histogram of the total number of steps taken each day, calculate and report the mean and median

4.1.Aggregte steps by date
```{r, echo = TRUE}
TotalsPerDay_I <- aggregate(ImputtedData$steps, list(ImputtedData$date), FUN=sum)
colnames(TotalsPerDay_I) <- c("Date", "Steps")
head(TotalsPerDay_I, n=20)
```

4.2.Plot histogram of the total number of steps taken each day
```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)

g <- ggplot(TotalsPerDay_I, aes(Steps))
g <- g + geom_histogram(boundary=0, binwidth=1000,col="black", fill="lightblue") + 
         ggtitle("Total steps per day") + xlab("Steps") + ylab("Frequency")
print(g)
```
4.3.Calculate and report the mean
```{r, echo = TRUE}
as.integer(mean(TotalsPerDay_I$Steps, na.rm=TRUE))
```

4.4.Calculate and report the median
```{r, echo = TRUE}
as.integer(median(TotalsPerDay_I$Steps, na.rm=TRUE))
```
Imputing the data resulted in a more normal distribution. As missing data was replaced with the mean values, the median moved and become equal to the mean.

## Differences in activity patterns between weekdays and weekends

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r, echo = TRUE}
ImputtedData$date <- as.Date(strptime(ImputtedData$date, format="%Y-%m-%d"))
ImputtedData$daytype <- sapply(ImputtedData$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
head(ImputtedData) 
```
2.Make a panel plot containing of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r, echo = TRUE}
plotdata <- aggregate(steps~interval + daytype, ImputtedData, mean, na.rm = TRUE)
g<- ggplot(plotdata, aes(x = interval , y = steps, color = daytype)) +
       geom_line() +
       labs(title = "Average Number of steps taken by day type (weekday/weekend)", x = "Interval", y = "Number of steps") +
       facet_wrap(~daytype, ncol = 1, nrow=2)
print(g)
```


## -- // END // -- 





